\fontfamily{phv}\selectfont{Causal Discovery methods for Effective Connectivity}

### 2013-12-09 docs/paper/paper.tex
plaatje van structuur max toegevoegd. Ook nog een scriptje toegevoegd om netwerk-filetjes te maken die met netDraw te openen zijn, maar die lijkt daar nog onzin van te maken (ligt naar mijn weten aan netDraw)

% Needed packages
\documentclass[a4paper, 10pt, english, onecolumn]{article}
\usepackage[english]{babel}
%\usepackage[cm]{fullpage}
\usepackage{cite}
\usepackage{anysize}
%\usepackage[compact]{titlesec}
\usepackage{graphicx}
\usepackage{stfloats}
\usepackage{listings}
\usepackage{hyperref}

@@

\setlength{\parindent}{0.0cm}
\setlength{\parskip}{0.25cm}

@@

\subsection{Pseudocode}
\label{sec:pseudocode}
The first part of the algorithm finds an undirected graph $G$. For each pair of points which are not connected in this graph, it also finds a separating set. $\operatorname{Adjacencies} (X)$ is defined as the set of all neighbours of $X$ in the graph $G$. This graph changes during the algorithm and so does $\operatorname{Adjacencies} (X)$.

\begin{algorithm}
\caption{Structure PC}
\begin{algorithmic}
\Function{Structure\_PC}{$data$}
\State connect all pairs of point in undirected graph $G$.
\State $n \gets 0$
\While{there are connected pairs $X$, $Y$ such that $n < |\operatorname{Adjecencies}(X)\setminus\{Y\}| $}
  \ForAll{such connected pairs $X$, $Y$}
    \ForAll{subsets $S$ of $\operatorname{Adjacencies(X)}$ for which $|S| = n$}
      \Comment{(Fisher-Z-test on $data$)}
      \State test whether $X$ and $Y$ are d-separated given $S$
      \If{$X$ and $Y$ are d-separated given $S$}
        \State remove the edge $X - Y$ from G
        \State record $S$ as a separating set of $X$ and $Y$
      \EndIf
    \EndFor
  \EndFor
  \State $n \gets n+1$
\EndWhile
\Return $G$ and the separating sets
\EndFunction
\end{algorithmic}
\label{code:emsstructure}
\end{algorithm}

The second part orients edges in this graph $G$ based on the separating sets.
\begin{algorithm}
\caption{Directional PC}
\begin{algorithmic}
\Function{Directional\_PC}{$G$, $sepsets$}
  \State $\text{PDAG} \gets G\text{, with all connections replaced by }\oline$
  \ForAll{unshielded triples $<X, Y, Z>$}
    \If{$Y$ is not in any separating set of $X$ and $Z \And X$ and $Z$ are dependent given $Y$}
      \Comment{V-structure-rule}
      \State Orient $X \sline Y \sline Z$ as $X \srightarrow Y \sleftarrow Z$ in $\text{PDAG}$
    \EndIf
  \EndFor
  \Repeat
    \ForAll{ordered, connected pairs $X$, $Y$}
      \Comment{Rule 1}
      \If{$X \osline Y$ or $X \sleftline Y$, and there is a node $Z$ such that $Z \srightarrow X$}
        \State Orient $X \sline Y$ as $X \rightarrow Y$
      \EndIf
      \Comment{Rule 2}
      \If{there is a directed path from $X$ to $Y$}
        \State Orient $X \sline Y$ as $X \srightarrow Y$
      \EndIf
    \EndFor
  \Until{no more edges are oriented}
  \State \Return PDAG
\EndFunction
\end{algorithmic}
\label{code:emsdirectional}
\end{algorithm}

@@

The results from the altorithms are presented in the form of matrices.
In such a matrix, indices 1-45 represent the brain regions of the left hemisphere, regions 45-90 represent the right hemisphere, and regions 91-116 represent regions in the cerebellum. %TODO left or right hemisphere?

Modified PC and EMS-PC do not differ in finding the structure.
The results are presented in figure ~\ref{fig:struct_full_mean} as the avarage over all six subjects.
\begin{figure}[h!]
  \centering
  \includegraphics{images/struct_full_mean_gray}
  \caption{The avarage structure of all subjects}
  \label{fig:struct_full_mean}
\end{figure}
%TODO aantal iteraties
In this matrix, higher values represent strong certainties of the presence of a connection.

The regions which are most clearly connected are the two hemispheres, and certain regions within the cerebellum.

>>>>

% Needed packages
\documentclass[a4paper, 10pt, english, onecolumn]{article}
\usepackage[english]{babel}
%\usepackage[cm]{fullpage}
\usepackage{cite}
%\usepackage{anysize}
%\usepackage[compact]{titlesec}
\usepackage{graphicx}
\usepackage{stfloats}
\usepackage{listings}
\usepackage{hyperref}

@@

%\setlength{\parindent}{0.0cm}
%\setlength{\parskip}{0.25cm}

@@

\subsection{Pseudocode}
\label{sec:pseudocode}

The first part of the algorithm finds an undirected graph $G$ and is described in Algorithm ~\ref{code:emsstructure}.
For each pair of points which are not connected in this graph, it also finds a separating set. $\operatorname{Adjacencies} (X)$ is defined as the set of all neighbours of $X$ in the graph $G$.
This graph changes during the algorithm and so does $\operatorname{Adjacencies} (X)$.
The second part orients edges in this graph $G$ based on the separating sets, as described in Algorithm ~\ref{code:emsdirectional}.

%\par\vfill\break % Omdat pseudocode veel hspace nodig heeft :(
%\advance\hsize by 8cm % Advance page height
%\advance\hoffset by -4cm % Shift top margin
%\floatevery{algorithm}{\setlength\hsize{15cm}}
%\floatevery{algorithmic}{\setlength\hsize{15cm}}
%TODO: meer hspace geven aan algorithms, alles hierboven werkte niet
\begin{algorithm}
\caption{Structure EMS-PC}
\begin{algorithmic}
\Function{Structure\_PC}{$data$}
\State connect all pairs of point in undirected graph $G$.
\State $n \gets 0$
\While{there are connected pairs $X$, $Y$ such that $n < |\operatorname{Adjecencies}(X)\setminus\{Y\}| $}
  \ForAll{such connected pairs $X$, $Y$}
    \ForAll{subsets $S$ of $\operatorname{Adjacencies(X)}$ for which $|S| = n$}
      \State test whether $X$ and $Y$ are d-separated given $S$
      \Comment{(Fisher-Z-test on $data$)}
      \If{$X$ and $Y$ are d-separated given $S$}
        \State remove the edge $X - Y$ from G
        \State record $S$ as a separating set of $X$ and $Y$
      \EndIf
    \EndFor
  \EndFor
  \State $n \gets n+1$
\EndWhile
\Return $G$ and the separating sets
\EndFunction
\end{algorithmic}
\label{code:emsstructure}
\end{algorithm}

\begin{algorithm}
\caption{Directional EMS-PC}
\begin{algorithmic}
\Function{Directional\_PC}{$G$, $sepsets$}
  \State $\text{PDAG} \gets G\text{, with all connections replaced by }\oline$
  \ForAll{unshielded triples $<X, Y, Z>$}
    \If{$Y$ is not in any separating set of $X$ and $Z \And X$ and $Z$ are dependent given $Y$}
      \Comment{V-structure-rule}
      \State Orient $X \sline Y \sline Z$ as $X \srightarrow Y \sleftarrow Z$ in $\text{PDAG}$
    \EndIf
  \EndFor
  \Repeat
    \ForAll{ordered, connected pairs $X$, $Y$}
      \Comment{Rule 1}
      \If{$X \osline Y$ or $X \sleftline Y$, and there is a node $Z$ such that $Z \srightarrow X$}
        \State Orient $X \sline Y$ as $X \rightarrow Y$
      \EndIf
      \Comment{Rule 2}
      \If{there is a directed path from $X$ to $Y$}
        \State Orient $X \sline Y$ as $X \srightarrow Y$
      \EndIf
    \EndFor
  \Until{no more edges are oriented}
  \State \Return PDAG
\EndFunction
\end{algorithmic}
\label{code:emsdirectional}
\end{algorithm}

%\par\vfill\break % Omdat pseudocode veel hspace nodig heeft :(
%\advance\hsize by -8cm % Retur n old margings and page height
%\advance\hoffset by 4cm % Return old margings and page height

@@

The results from the altorithms are presented in the form of matrices.
A connection from region $n$ to region $m$ is represented by point $(n,m)$ or $n,m$ in the matrix.
For strucural data, this matrix is therefore symmetrical.
In the directional data, point $(n,m)$ denotes a directed edge from $n$ to $m$.
In such a matrix, indices 1-45 represent the brain regions of the left hemisphere, regions 45-90 represent the right hemisphere, and regions 91-116 represent regions in the cerebellum. %TODO left or right hemisphere?

Modified PC and EMS-PC do not differ in finding the structure.
The results are presented in figure ~\ref{fig:struct_full_mean} as the avarage over all six subjects, with 300 iterations per subject.
\begin{figure}[h!]
  \centering
  \includegraphics{images/struct_full_mean_gray}
  \caption{The avarage structure of all subjects}
  \label{fig:struct_full_mean}
\end{figure}
In this matrix, higher values represent strong certainties of the presence of a connection.

\begin{figure}[h!]
  \centering
  \includegraphics{images/structure_max}
  \caption{The avarage structure of all subjects, according to the DWI-measurements of Hinne et al.}
  \label{fig:struct_mean_max}
\end{figure}

One notable difference between the results found in Hinne et al. are that connections between the the two hemispheres are clearly visible as two diagonal lines. Connections within the hemispheres are less clearly visible, as well as connections within the cerebellum.
